---
title: "INFSCI 2809: Spatial Data Analytics - Final Exam"
output: html_notebook
---
Raanan Gurewitsch
April 22, 2019
```{r}
# Packages
library(plyr)
library(dplyr)
library(data.table)
library(rgdal)
library(spatstat)
library(GISTools)
library(maptools)
library(maps)
library(spdep)
library(spgwr)
library(rgeos)
library(ggplot2)
library(ape)
library(phylin)
library(raster)
library(data.table) 
library(mapproj)
library(PBSmapping)
library(SDraw)
library(tmap)
library(tmaptools)
library(sf)
```

```{r}


```
PART A: 
```{r}
neigh1 <- readOGR("C:/Users/raana/OneDrive/Desktop/GIS/DataFiles_FinalExam/1_Neighbor1","1_Neighbor1")
neigh2 <- readOGR("C:/Users/raana/OneDrive/Desktop/GIS/DataFiles_FinalExam/1_Neighbor2","1_Neighbor2")
```
Perform autocorrelation on the population attribute (POP_ARR02) in "1_Neighbor1.shp" and "1_Neighbor2.shp" files by using Rook's adjacency and Queen's adjacency.
```{r}
# Create queens and rooks adjacency matrices for both files
Wq1 <- poly2nb(neigh1)
Wr1 <- poly2nb(neigh1, queen = FALSE)

Wq2 <- poly2nb(neigh2)
Wr2 <- poly2nb(neigh2, queen = FALSE)
# Transform into spatial weights matrix
Wmq1 <- nb2mat(Wq1, style='B')
Wmr1 <- nb2mat(Wr1, style='B')

Wmq2 <- nb2mat(Wq2, style='B')
Wmr2 <- nb2mat(Wr2, style='B')
```
Use both Moran's I and Geary's C to measure the level, type, significance of correlation in the attribute.
```{r}
# Using Rook's adjacency, here we calculate the Moran's I value for both files
x1 <- neigh1$POP_ARR02
x2 <- neigh2$POP_ARR02
moran_r1 <- Moran.I(x1, Wmr1)
moran_r2 <- Moran.I(x2, Wmr2)
# Now using Queen's adjacency, here we calculate the Moran's I value for both files
moran_q1 <- Moran.I(x1, Wmq1)
moran_q2 <- Moran.I(x2, Wmq2)

print("File 1 Moran's:")
print("Rook's: ")
moran_r1
print("Queen's: ")
moran_q1
      
print("File 2 Moran's:")
print("Rook's: ")
moran_r2
print("Queen's: ")
moran_q2
```
```{r}
qtm(shp = neigh1, fill = "POP_ARR02", fill.palette = "-Reds", title = "Neighborhood 1") 
qtm(shp = neigh2, fill = "POP_ARR02", fill.palette = "Blues", ncol = 2, title = "Neighborhood 2")
```
```{r}
# Before performing the Geary's C test, we have to convert the spatial weights matrices into list form
Wmr1_l <- nb2listw(Wr1)
Wmr2_l <- nb2listw(Wr2)
Wmq1_l <- nb2listw(Wq1)
Wmq2_l <- nb2listw(Wq2)
# Using Rook's adjacency, here we calculate the Geary's C value for both files
geary_r1 <- geary.test(x1, Wmr1_l)
geary_r2 <- geary.test(x2, Wmr2_l)
# Now using Queen's adjacency, here we calculate the Geary's C value for both files
geary_q1 <- geary.test(x1, Wmq1_l)
geary_q2 <- geary.test(x2, Wmq2_l)

print("File 1 Geary's C:")
print("Rook's: ")
geary_r1
print("Queen's: ")
geary_q1
      
print("File 2 Geary's C:")
print("Rook's: ")
geary_r2
print("Queen's: ")
geary_q2
```

Part B: 
Use the Inverse Distance Weighting (IDW) and the Ordinary Kriging (OK) methods to
interpolate an "Intensity" value at the center location of each polygon in "2_Community.shp" file using the sample "Intensity" values at locations (s_lat, s_long) in the same file. 

For IDW, use different values for r (window size) and different values for k (exponent of distance) to find the most similar map as produced by OK. Use exponential semiveriogram model for OK.

```{r}
Comm2 <- readOGR("C:/Users/raana/OneDrive/Desktop/GIS/DataFiles_FinalExam/2_Community","2_Community")

# Get Interpolated surface using Ordinary Kriging

# Set up grid system for interpolation
bbox(Comm2)
proj4string(Comm2)
coords_ = coordinates(Comm2)
grid_ <- data.frame(x=coords_[,1], y=coords_[,2])
coordinates(grid_) = ~x+y

# Generate exponential semivariogram 
vgmodel <- variogram(Intensity~1, Comm2)
vgfit <- fit.variogram(vgmodel, model = vgm(1, "Exp",12,1))
proj4string(Comm2) = CRS(proj4string(grid_))
intld <- krige(Intensity~1, Comm2, grid_, model=vgfit)

# Add column with interpolated values
Comm2$Interpolated <- intld$var1.pred

# Generate map of interpolated surface (OK)
qtm(Comm2, fill = "Interpolated", fill.palette = "Blues", title = 'Interpolated Values (OK)') +
tm_scale_bar(size = 0.5) + tm_grid(n.x = 5, n.y = 5) +
tm_compass(position = c('right', 'top'))

```

```{r}
# Test 4 different powers for idw interpolation
test2.idw <- gstat::idw(Intensity~1, Comm2, newdata=grid_,idp=2)
test4.idw <- gstat::idw(Intensity~1, Comm2, newdata=grid_,idp=4)
test6.idw <- gstat::idw(Intensity~1, Comm2, newdata=grid_,idp=6)
test8.idw <- gstat::idw(Intensity~1, Comm2, newdata=grid_,idp=8)
print(test2.idw$var1.pred)
print(test4.idw$var1.pred)
print(test6.idw$var1.pred)
print(test8.idw$var1.pred)

#[inverse distance weighted interpolation]
#[inverse distance weighted interpolation]
#[inverse distance weighted interpolation]
#[inverse distance weighted interpolation]
# [1]   8  64  20 150  14   5 200  75  50  18   7 175
# [1]   8  64  20 150  14   5 200  75  50  18   7 175
# [1]   8  64  20 150  14   5 200  75  50  18   7 175
# [1]   8  64  20 150  14   5 200  75  50  18   7 175

# Test shows that each power produces same results


```

```{r}

# Add column with interpolated values
Comm2$Interpolated_IDW <- test2.idw$var1.pred

# Generate map of interpolated surface (OK)
qtm(Comm2, fill = "Interpolated_IDW", fill.palette = "Blues", title = 'Interpolated Values (IDW)') +
tm_scale_bar(size = 0.5) + tm_grid(n.x = 5, n.y = 5) +
tm_compass(position = c('right', 'top'))



```

```{r}
# Get interpolated surface using IDW 
test_idw <- idw(Comm2, coords_, grid_, p = 2, R = 2)


# Calculate interpolated values using different powers

for (p_ in 2:8){
  
  for (r_ in 2:8){
    # Generate interpolated surface 
    test_idw <- idw(Comm2, coords_, grid_, p = i, R = j)
    
    
  }
  
  
}





# Calculate interpolated values using different window sizes
#ipd_p2 <- idw(Comm2, coords_, grid_, p = 2)
#ipd_p4 <- idw(Comm2, coords_, grid_, p = 4)
#ipd_p6 <- idw(Comm2, coords_, grid_, p = 6)
#ipd_p8 <- idw(Comm2, coords_, grid_, p = 8)


ipd_p2
grid.image(ipd_p2, grid_, main="IDW Interpolated Surface",xlab = "Longitude", ylab = "Latitude")
points(grid_)
```

```{r}
Comm2
```

```{r}

```

```{r}

```

```{r}

```
Part C: 
```{r}
P1 <- readOGR("C:/Users/raana/OneDrive/Desktop/GIS/DataFiles_FinalExam/3_Area1", "3_Area1")
P2 <- readOGR("C:/Users/raana/OneDrive/Desktop/GIS/DataFiles_FinalExam/3_Area2", "3_Area2")
P3 <- readOGR("C:/Users/raana/OneDrive/Desktop/GIS/DataFiles_FinalExam/3_TrailPoints","3_TrailPoints")
P4 <- readOGR("C:/Users/raana/OneDrive/Desktop/GIS/DataFiles_FinalExam/3_State_Roads","3_State_Roads")
```
a. Overlay P2 on P1 and find their intersection (P12), P1/(=P1-P12), and P2/(=P2-P12).
```{r}
a1 <- combinePolys(SpatialPolygons2PolySet(P1))
a2 <- combinePolys(SpatialPolygons2PolySet(P2))
# Get intersection P12 
P12 <- combinePolys(joinPolys(a1, a2,"INT"))
plotPolys(a12, proj = TRUE, col="blue2",xlab="Longitude",ylab="Latitude")
title("P12")
# Get P1' by subtracting P1 and P12
P1_ <- combinePolys(joinPolys(a1, a12,"DIFF"))
plotPolys(P1_, proj = TRUE, col="darkorange1",xlab="Longitude",ylab="Latitude")
title("P1'")
# Get P2' by subtracting P2 and P12
P2_ <- combinePolys(joinPolys(a2, a12,"DIFF"))
plotPolys(P2_, proj = TRUE, col="gold2",xlab="Longitude",ylab="Latitude")
title("P2'")
```
b. Overlay P3 on P1, P2, and P12 and find number of points within P1, P2, and P12.
```{r}
# For P1:
plotPolys(a1, proj = TRUE, col=alpha("gold2",0.1),xlab="Longitude",ylab="Latitude")
title("P3 in P1")
proj4string(P3) <- proj4string(P1)
P3P1 <- !is.na(over(P3, as(P1, "SpatialPolygons")))
points(P3[P3P1, ], pch=16, col="red")
# For P2:
plotPolys(a2, proj = TRUE, col=alpha("gold2",0.1),xlab="Longitude",ylab="Latitude")
title("P3 in P2")
proj4string(P3) <- proj4string(P2)
P3P2 <- !is.na(over(P3, as(P2, "SpatialPolygons")))
points(P3[P3P2, ], pch=16, col="red")
# For P12:
plotPolys(a12, proj = TRUE, col=alpha("gold2",0.1),xlab="Longitude",ylab="Latitude")
title("P3 in P12")
P12_raster <- raster::intersect(P1, P2)
#class(P12_raster)
proj4string(P3) <- proj4string(P12_raster)
P3P12 <- !is.na(over(P3, as(P12_raster, "SpatialPolygons")))
points(P3[P3P12, ], pch=16, col="red")
```
Results:
```{r}
# Count the points in each region
print(paste0("Total points:",length(P3)))
print(paste0("      In P1: ",length(P3[P3P1, ])))
print(paste0("      In P2: ",length(P3[P3P2, ])))
print(paste0("      In P12: ",length(P3[P3P12, ])))
```
c. Overlay P4 on P1, P2, and P12 and find the total length of road segments within P1, P2, and
P12
```{r}
plotPolys(a1, proj = TRUE, col=alpha("gold2",0.1),xlab="Longitude",ylab="Latitude")
title("P4 in P1")
proj4string(P4) <- proj4string(P1)
P4P1 <- !is.na(over(P4, as(P1, "SpatialPolygons")))
plot(P4[P4P1, ],add=T, col="red")

plotPolys(a2, proj = TRUE, col=alpha("gold2",0.1),xlab="Longitude",ylab="Latitude")
title("P4 in P2")
proj4string(P4) <- proj4string(P2)
P4P2 <- !is.na(over(P4, as(P2, "SpatialPolygons")))
plot(P4[P4P2, ],add=T, col="red")
```
```{r}
plotPolys(a12, proj = TRUE, col=alpha("gold2",0.1),xlab="Longitude",ylab="Latitude")
title("P4 in P12")
proj4string(P4) <- proj4string(P12_raster)
P4P12 <- !is.na(over(P4, as(P12_raster, "SpatialPolygons")))
plot(P4[P4P12, ],add=T, col="red")

```
```{r}
calcArea (a12, rollup = 1)

print(paste0("Total road segment length: ",sum(P4$SEG_LNGTH_)))
print(paste0("                   in  P1: ",lineLength(P4[P4P1,],byid=FALSE)))
print(paste0("                   in  P2: ",lineLength(P4[P4P2,],byid=FALSE)))
print(paste0("                   in  P12: ",lineLength(P4[P4P12,],byid=FALSE)))
```

```{r}

```























